The role of cilia-mediated cerebrospinal fluid flow in intraventricular transport in zebrafish embryos

### 2025-01-27 paper/main.tex
update paper, clean utilities

% Title
\title{Cilia-mediated flow and transport in zebrafish brain ventricles in vivo and in silico}
\author[1,*]{Halvor Herlyng}
\author[1]{Ada J.~Ellingsrud}
\author[1]{Miroslav Kuchta} 
\author[X]{Inyoung Jeong} % OK?
\author[1,3,+]{Marie E.~Rognes} % OK order?
\author[2,+]{Nathalie Jurisch-Yaksi}
\affil[1]{Department for Numerical Analysis and Scientific Computing, Simula Research Laboratory, Oslo, Norway}
\affil[2]{Department of Clinical and Molecular Medicine, Norwegian University of Science and Technology, Trondheim, Norway}
\affil[3]{K.~G.~Jebsen Center for Brain Fluid Research}
\affil[*]{\href{mailto:hherlyng@simula.no}{hherlyng@simula.no}}
\affil[+]{these authors contributed equally to this work}
% Optional PDF information
\ifpdf
\hypersetup{
  pdftitle={cilia-csf},
  pdfauthor={H.~Herlyng \textit{et al.}}
}
\fi

\begin{abstract}
    This paper presents ...
    %Flow of cerebrospinal fluid in zebrafish brain ventricles is studied in this work. Motivated by experimental observations, we consider motile cilia and cardiac beating as the driving flow mechanisms. A mathematical model of the flow is developed and coupled with a transport model to study the transport of molecules inside the brain ventricles. The Stokes equations are used to model the flow, with the motile cilia acting as a force on the fluid through a stress boundary condition that allows slip at the wall. Transport is modeled with an advection-diffusion equation coupled to the velocity field obtained from solving the Stokes equations. The flow and transport models are discretized with finite element methods. For the Stokes equations we employ a non-conforming discretization based on elements of the Brezzi-Douglas-Marini family. The advection-diffusion equation is discretized with discontinuous Galerkin elements. We implement and solve the discretized equations numerically to simulate flow and transport in a zebrafish brain ventricles geometry obtained with medical imaging.
\end{abstract}
\begin{document}
\maketitle
% Introduction:
% 1. What are cilia and why are they important?
% 2. Previous studies of cilia flow
% 3. What problem do we want to solve
% 4. The model and work done in this paper
\section*{Introduction}
Development of brains and sustaining healthy functioning of the brain relies on a myriad of regulatory processes. One of these processes is the flow of cerebrospinal fluid (CSF) in the brain ventricles. The brain ventricles are hollow structures inside the skull where CSF is produced and flows around, transporting nutrients, neurotransmitters and other molecules important for neural development.

@@

\subsection*{Intraventricular injection of microbeads and particle tracking}
In addition to photoconversion experiments, we carried out particle tracking inside the brain ventricles of a larval zebrafish, to demonstrate the motile-cilia and cardiac-cycle related flow features observed and described in Olstad~\emph{et al.}~\cite{Olstad2019CiliaryDevelopment}. A schematical illustration of the flow features in the brain ventricles is shown in~\cref{fig:figure1}, panel \textbf{a}. The experimental setup is identical (\lyng{HH: Is this true?}) to the ones considered there, and we refer to that paper for further details than the ones provided here in the following.

@@

\subsection*{Solute transport}
We simulate transport of photoconverted proteins in the brain ventricles by modeling transport of a solute with concentration $c(\xx, t)$ for $\xx\in\Omega$ and time $t > 0$ by the advection-diffusion equation
\begin{equation}
    \pdifft{c} + \nabla\cdot\JJ = 0\mbox{  in }\Omega,
    \label{eq:adv_diff_strong}
\end{equation}
where $\mathbf{J}$ is the total concentration flux. The total concentration flux is assumed to consist of an advective flux and a diffusive flux:
\begin{equation*}
    \JJ = c\uu - D\nabla c\mbox{ in }\Omega.
\end{equation*}
Here, $\uu$ is the cerebrospinal fluid velocity field, governed by the Stokes equations~\eqref{eq:stokes_eqs}, and $D$ is the molecular diffusion coefficient of the solute. The diffusion coefficient was estimated through extrapolation of previous measurements~\cite{GuraSadovsky2017MeasurementExpansion, Mahmood2023ExosomeTemperature, Swaminathan1997PhotobleachingDiffusion, Potma2001ReducedCells} via the Stokes-Einstein relation. We elaborate on the extrapolation in a later section.

@@

\subsection*{Physical parameters and dimensionless numbers \lyng{HH: Change this to "quantities of interest" and add mean concentration and time to threshold?}}\label{subsec:parameters}
The cerebrospinal fluid density $\rho$ and dynamic viscosity $\mu$ were set following the work of Bloomfield \textit{et al.}~\cite{Bloomfield1998EffectsFluid}~(\cref{tab:parameter_table}). We considered a range of small, medium and large molecules with three different diffusion coefficients $D_1, D_2$ and $D_3$ (cf. the following section for details on determining their values). Experimental data~\cite{Olstad2019CiliaryDevelopment} indicate velocities proximate to the ventricular walls of $27.4 \pm 5.4 \ \mu$m/s and $2.6 \pm 0.6 \ \mu$m/s in the dorsal and ventral regions of the diencephalic ventricle, respectively. In the telencephalic ventricle, the velocities were estimated at $4.7 \pm 1.3 \ \mu$m/s. We calibrated $\tau$ and $A$ heuristically by comparing these experimental velocity values with the simulated, maximum velocity magnitude. In addition, particle tracking data of pulsatile motion~(\cref{sec:appC_particle_transport}) was used in the calibration process to determine the relative contributions of the cardiac pulsations, as compared to the cilia-driven flow, to the velocity magnitude.
\begin{table}[h!]
    \centering
    \caption{Physical parameters of the flow and transport models. (*) Calibrated based on experimental data~\cite{Olstad2019CiliaryDevelopment}. (\textdagger) Estimated based on experimental data~\cite{GuraSadovsky2017MeasurementExpansion, Swaminathan1997PhotobleachingDiffusion, Potma2001ReducedCells}.}
    \label{tab:parameter_table}
    \begin{tabular}{lcccc}
    \toprule
        Parameter & Symbol & Value & Unit & Reference \\
        \midrule
        Fluid density & $\rho$ & $1000$ & $\mathrm{kg/m^3}$ & \cite{Bloomfield1998EffectsFluid} \\
        Dynamic viscosity & $\mu$ & $\num{6.97e-04}$ & $\mathrm{Pa\cdot s}$ & \cite{Bloomfield1998EffectsFluid} \\
        Cardiac frequency & $f$ & 2.22 & Hz &  \cite{Olstad2019CiliaryDevelopment}\\
        Angular cardiac frequency & $\omega = 2\pi f$ & 6.97 & rad/s & \cite{Olstad2019CiliaryDevelopment}\\
        Cilia tangential traction & $\tau$ & \num{6.5e-04} & Pa & * \\
        Pressure amplitude & $A$ & \num{1.5e-03} & Pa & * \\
        %Ref. frame motion amplitude & $A_{\mathrm{ref}}$ & \num{9.19e-04} & m/s & *\\
        Exosome diffusion coefficient & $D_1$ & $\num{1.63e-12}$ & $\mathrm{m^2/s}$ & \textdagger\\
        STM+GFP diffusion coefficient  & $D_2$ & $\num{5.75e-11}$ & $\mathrm{m^2/s}$ & \textdagger\\
        Dendra2 diffusion coefficient  & $D_3$ & $\num{1.15e-10}$ & $\mathrm{m^2/s}$ & \textdagger\\
    \bottomrule
    \end{tabular}
\end{table}

@@

\subsection*{Model versions and physiological scenarios}\label{subsec:model_versions}
We have introduced boundary conditions to model two separate driving forces of cerebrospinal fluid flow, namely motile cilia and cardiac pulsations. To distinguish the impact of each component, three different setups for the flow model were considered: 
\begin{itemize}
    \item Model 0: The baseline model representing a wild-type zebrafish. Both cilia forces and cardiac pulsations are included in the model.

@@

    %\item Model IIb: Only cardiac pulsations are modeled, using the domain movement approach. No cilia forces.
\end{itemize}
When simulating transport of solute in the ventricles, we consider the same six different regions of interest that were considered in the photoconversion experiments. In region 1, we use the function~\eqref{eq:injection_curve} to enforce the concentration such that it represents the experimental data of photoconversion of proteins. We consider the three different diffusion coefficients introduced in previous sections to simulate and compare results for transport of molecules of varying sizes.

\lyng{Add information on removing cilia and modifying geometries simulation studies in this section?}
\begin{table}[h!]
    \centering
    \caption{A summary of the flow model versions. For all model versions, we consider a free slip boundary condition on $\Gs=\Gamma\setminus\left(\Gc\cup\Gp\right)$.}
    \label{tab:model_versions}
    \begin{tabular}{c|c@{\hskip 0.4in}c}
       \toprule
       Model & Driving force & Flow BCs \\ %& Domain motion \\
       \midrule
       Baseline (0)   & Cilia and cardiac & \makecell[l]{Impermeability $\uu\cdot\nn=0$ and tangential traction $\bsigpar=\btau$ on $\Gc$, \\ normal pressure $\Tilde{p}(t)$ on $\Gp$} \\[0.3in] %& No\\[0.3in]
       No cardiac (I) & Cilia   & \makecell[l]{Impermeability $\uu\cdot\nn=0$ and tangential traction $\bsigpar=\btau$ on $\Gc$} \\[0.3in]%& No\\[0.3in]
       No cilia (II) & Cardiac & \makecell[c]{Normal pressure $\Tilde{p}(t)$ on $\Gp$} \\% & No\\[0.3in]
       %No cilia (IIb) & Cardiac & \makecell[l]{Impermeability $\mathbf{w}\cdot\nn=0$ and \\ free slip $\bsigpar(\mathbf{w}, p)=\mathbf{0}$ on $\Gamma$}  & Yes \\
       \bottomrule
    \end{tabular}
\end{table}

@@

In the absence of pressure boundary conditions, the pressure in the Stokes equations is only determined up to an additive constant. Thus, with the boundary conditions used in model version I, the pressure is not uniquely defined and discretization of the Stokes equations leads to a singular linear equation system. \lyng{To make the discrete problem well-posed for this setup, the right-hand side vector of the linear system assembled after discretization of the Stokes equations equations is orthogonalized with respect to the basis of the nullspace of the pressure to ensure existence of solutions. This nullspace is the set of (real) scalar constants. Furthermore, information of the pressure nullspace is provided to the linear solver, such that a unique solution is determined.}

@@

%%--------------RESULTS--------------%%
\section*{Results}\label{sec:results}
\subsection*{Cerebrospinal fluid flow in the brain ventricles is compartmentalized}
With cilia forces and cardiac forces represented in the flow model (model version 0), we observe a vortex-structured flow pattern in the anterior and middle ventricles~(\cref{fig:figure1}, panel \textbf{g} \lyng{Need to reorder panels in figure 1, since results for model 0 are introduced first.}). This rotational motion shows that cilia-mediated flow is dominating in these regions, as the vortex structures are a consequence of the applied tangential stresses on the ciliated boundaries combined with the assumption of fluid incompressibility. In the posterior ventricle, we observe directional flow as the pressure-driven flow is dominating in this region. The cardiac pressure forces induce a pulsatile motion of the cerebrospinal fluid synchronized with the cardiac cycle, with velocity magnitudes of the pulsatile motion being approximately an order of magnitude smaller than the velocities related to the cilia-driven motion. Over a cardiac cycle, the maximum velocity magnitude in the ventricles was 27.7 $\mu$m/s, located dorsally in the posterior part of the middle ventricle. This maximum velocity was attained at the time instant coinciding with the peak of the cardiac pulsatile flow flowing caudally, as a result of both the tangential stress boundary condition and the pulsatile pressure-driven flow being aligned with the rostrocaudal axis (the $x$ axis). A calculation using the observed maximum velocity of 27.7 $\mu$m/s, and choosing the approximate height of the middle ventricle $110 \ \mu$m as a characteristic length scale, yields a Reynolds number $\mathrm{Re} \approx 0.004$, justifying the assumption of Stokes flow.

\subsection*{Vortex structures vanish in the absence of cilia}
Excluding the cilia tangential traction from the model (model II), vortex-structured compartmentalization of the ventricular flow does not occur, and the flow is directional and pulsatile as a result of the pressure boundary conditions imposed~(\cref{fig:figure1}, panel \textbf{d}). The pulsatile motion is absent when disregarding cardiac motion~(model I, see \cref{fig:figure1}, panel \textbf{d}). For model I, the boundary conditions are not time-dependent, rendering the solution to the Stokes equations~\eqref{eq:stokes_eqs} a steady-state flow field. Streamlines demonstrate the compartmentalization of the cerebrospinal fluid flow driven by cilia, and we observe highest vorticity in the vicinity of cilia.

@@

We quantify the transport of solute by showing the profile of the mean concentrations
\begin{equation}
    \overline{c}_i = \frac{1}{|\Omega_i|}\int_{\Omega_i} c\dx,
    \label{eq:c_mean_i}
\end{equation}
in the regions of interest (ROIs) $\Omega_i$ as functions of time, where $|\Omega_i|$ is the volume of the domain $\Omega_i$. The $\overline{c}_i$ profiles and an illustration of the ROIs are presented in panels~\textbf{b--e} of~\cref{fig:figure2_compare_diffusion_coefficients}. The dynamics of ROI 1 follows from imposing~\cref{eq:injection_curve}. Note that the total mass added to the system is not exactly the same for all three diffusion coefficients, since the total mass added depends on how fast the mass added in ROI 1 is transported away from that region.

For ROIs 2--6, we observe more rapid concentration increases when the diffusion coefficient increases.  For all three diffusion coefficients, the dynamics of ROI 6 is the slowest. Region 5 exhibits a more rapid increase in mean concentration, and an even faster increase is observed for ROIs 2--4. The profiles of the latter three regions are similar, displaying only small differences in all of the three plots.

As another way of quantifying transport, we report the times $\hat{t}$ when the mean concentrations $\overline{c}_i$ first exceed a threshold value $\hat{c}$~(\cref{fig:figure2_compare_diffusion_coefficients}, panel \textbf{f}). We use $\hat{c}=0.25$ for regions 1--4 and $\hat{c}=0.10$ for regions 5 and 6. The choice of different threshold values is because of the slower dynamics in ROIs 5 and 6. In region 6, the threshold is never exceeded for $D_1$, with the mean concentration close to zero at the end of the simulation.

Finally, we calculate Péclet numbers for the entire ventricular geometry using estimates of the diffusion and advection timescales $t_d=L_p^2/D$ and $t_a=L_p/U_p$. Recall that
\begin{equation*}
    \mathrm{Pe} = \frac{\mathrm{diffusion \ timescale}}{\mathrm{advection \ timescale}} = \frac{t_d}{t_a},
\end{equation*}
so that if $\mathrm{Pe} > 1$, transport is dominated by advection, and if $\mathrm{Pe} < 1$ transport is dominated by diffusion. For the length scale $L_p$ we choose the approximate length of the ventricles geometry in the $x$-direction, $L_p=600 \ \mu$m. We use a mean velocity $\overline{U}$ as the characteristic velocity $U_p$. The mean velocity $\overline{U}$ is calculated by first averaging the velocity in space:
\begin{equation*}
    \overline{u}^2 = \frac{1}{|\Omega|}\int_{\Omega}\uu\cdot\uu\dx,
\end{equation*}
and then averaging $\overline{u}$ in time over one cardiac cycle to obtain $\overline{U}=1.41 \ \mu$m/s. This yields the three Péclet numbers $\mathrm{Pe}_1=519$, $\mathrm{Pe}_2=14.7$ and $\mathrm{Pe}_3=7.4$ for $D_1$, $D_2$ and $D_3$, respectively.

\subsection*{Comparing experimental and computational results}
Transport of photoconverted proteins was quantified by the relative change in fluorescence intensity $\Delta F=(F(t)-F_0)/F_0$ over time~(\cref{fig:figure3_compare_exp_and_sim_control}, panel \textbf{a}). The solid lines are the mean values of $\Delta F$ for the fish cohort, with shaded regions covering one standard deviation. The first region of interest (ROI 1) is the laser-targeted region. We remind the reader that the fluorescence intensity change $\Delta F$ is scaled by the maximum fluorescence intensity $F_0$ in ROI 1 to ease comparison of the profiles.

Qualitatively, the fluorescence intensity change in region one resembles intensity dynamics observed in previous experiments studying the photoconversion properties of Dendra2 fluorescent proteins~\cite{Makarov2014Steady-stateDendra2}. The transport to the other regions can be separated into two groups: the fluorescence intensity change $\Delta F$ for ROIs 2--4 in the middle ventricle are all similar, whereas the more distal regions ROI 5 and 6 (anterior and posterior ventricle, respectively) exhibit a slower increase in $F$. In ROIs 2--4, the $\Delta F$ curves are concave, similar to ROI 1, while in ROIs 5 and 6 the curves are convex. Although the fluorescence intensity increases throughout all of the experiments, the maximum $\Delta F$ is not necessarily attained at the final time of the experiments, mostly due to signal noise. At the final time, the mean fluoresence intensities in regions 2--4 all exceed 60\% (63\%, 71\%, 67\%, respectively) of the final value $F_0$ in ROI 1, wheras ROIs 5 and 6 reach mean values of 29\% and 18\%, respectively.

>>>>

% Title
\title{The role of cilia-mediated cerebrospinal fluid flow in intraventricular transport in zebrafish embryos}
\author[1,*]{Halvor Herlyng}
\author[1]{Ada J.~Ellingsrud}
\author[1]{Miroslav Kuchta} 
\author[X]{Inyoung Jeong} % OK?
\author[1,3,+]{Marie E.~Rognes} % OK order?
\author[2,+]{Nathalie Jurisch-Yaksi}
\affil[1]{Department for Numerical Analysis and Scientific Computing, Simula Research Laboratory, Oslo, Norway}
\affil[2]{Department of Clinical and Molecular Medicine, Norwegian University of Science and Technology, Trondheim, Norway}
\affil[3]{K.~G.~Jebsen Center for Brain Fluid Research}
\affil[*]{\href{mailto:hherlyng@simula.no}{hherlyng@simula.no}}
\affil[+]{these authors contributed equally to this work}
% Optional PDF information
\ifpdf
\hypersetup{
  pdftitle={zebrafish-cilia-csf},
  pdfauthor={H.~Herlyng \textit{et al.}}
}
\fi

\begin{abstract}
    This paper presents ...
    %Flow of cerebrospinal fluid in zebrafish brain ventricles is studied in this work. Motivated by experimental observations, we consider motile cilia and cardiac beating as the driving flow mechanisms. A mathematical model of the flow is developed and coupled with a transport model to study the transport of molecules inside the brain ventricles. The Stokes equations are used to model the flow, with the motile cilia acting as a force on the fluid through a stress boundary condition that allows slip at the wall. Transport is modeled with an advection-diffusion equation coupled to the velocity field obtained from solving the Stokes equations. The flow and transport models are discretized with finite element methods. For the Stokes equations we employ a non-conforming discretization based on elements of the Brezzi-Douglas-Marini family. The advection-diffusion equation is discretized with discontinuous Galerkin elements. We implement and solve the discretized equations numerically to simulate flow and transport in a zebrafish brain ventricles geometry obtained with medical imaging.
\end{abstract}
\begin{document}
\maketitle
% Introduction:
% 1. What are cilia and why are they important?
% 2. Previous studies of cilia flow
% 3. What problem do we want to solve
% 4. The model and work done in this paper
\section*{Introduction}
\lyng{HH: This needs work, have just jotted down some stuff to have references ready.}

Development of brains and sustaining healthy functioning of the brain relies on a myriad of regulatory processes. One of these processes is the flow of cerebrospinal fluid (CSF) in the brain ventricles. The brain ventricles are hollow structures inside the skull where CSF is produced and flows around, transporting nutrients, neurotransmitters and other molecules important for neural development.

@@

\subsection*{Intraventricular injection of microbeads and particle tracking}
In addition to photoconversion experiments, we carried out particle tracking inside the brain ventricles of a larval zebrafish, to demonstrate the motile-cilia and cardiac-cycle related flow features observed and described in Olstad~\emph{et al.}~\cite{Olstad2019CiliaryDevelopment}. A schematical illustration of the flow features in the brain ventricles is shown in~\cref{fig:figure1}, panel \textbf{a}. Our experimental setup for the particle tracking is identical to the ones considered there (\lyng{HH: Is this true?}), and we refer to that paper for further details than the ones provided here in the following.

@@

\subsection*{Computational model of solute transport}
We simulate transport of photoconverted proteins in the brain ventricles by modeling transport of a solute with concentration $c(\xx, t)$ for $\xx\in\Omega$ and time $t > 0$ by the advection-diffusion equation
\begin{equation}
    \pdifft{c} + \nabla\cdot\JJ = 0\mbox{  in }\Omega,
    \label{eq:adv_diff_strong}
\end{equation}
where $\mathbf{J}$ is the total concentration flux. The total concentration flux is assumed to consist of an advective flux and a diffusive flux:
\begin{equation*}
    \JJ = c\uu - D\nabla c\mbox{ in }\Omega.
\end{equation*}
Here, $\uu$ is the cerebrospinal fluid velocity field, governed by the Stokes equations~\eqref{eq:stokes_eqs}, and $D$ is the molecular diffusion coefficient of the solute. The diffusion coefficient was estimated through extrapolation of previous measurements~\cite{GuraSadovsky2017MeasurementExpansion, Mahmood2023ExosomeTemperature, Swaminathan1997PhotobleachingDiffusion, Potma2001ReducedCells} via the Stokes-Einstein relation. We elaborate on the extrapolation in a later section.

@@

\subsection*{Physical parameters}\label{subsec:parameters}
The cerebrospinal fluid density $\rho$ and dynamic viscosity $\mu$ were set following the work of Bloomfield \textit{et al.}~\cite{Bloomfield1998EffectsFluid}~(\cref{tab:parameter_table}). We considered a range of small, medium and large molecules with three different diffusion coefficients $D_1, D_2$ and $D_3$ (cf. the following section for details on determining their values). Experimental data~\cite{Olstad2019CiliaryDevelopment} indicate velocities proximate to the ventricular walls of $27.4 \pm 5.4 \ \mu$m/s and $2.6 \pm 0.6 \ \mu$m/s in the dorsal and ventral regions of the diencephalic ventricle, respectively. In the telencephalic ventricle, the velocities were estimated at $4.7 \pm 1.3 \ \mu$m/s. We calibrated $\tau$ and $A$ heuristically by comparing these experimental velocity values with the simulated, maximum velocity magnitude. In addition, particle tracking data of pulsatile motion~(\cref{sec:appC_particle_transport}) was used in the calibration process to determine the relative contributions of the cardiac pulsations, as compared to the cilia-driven flow, to the velocity magnitude.
\begin{table}[h!]
    \centering
    \caption{Physical parameters of the flow and transport models. (*) Calibrated based on experimental data~\cite{Olstad2019CiliaryDevelopment}. (\textdagger) Estimated based on experimental data~\cite{GuraSadovsky2017MeasurementExpansion, Swaminathan1997PhotobleachingDiffusion, Potma2001ReducedCells}.}
    \label{tab:parameter_table}
    \begin{tabular}{lcccc}
    \toprule
        Parameter & Symbol & Value & Unit & Reference \\
        \midrule
        Fluid density & $\rho$ & $1000$ & $\mathrm{kg/m^3}$ & \cite{Bloomfield1998EffectsFluid} \\
        Dynamic viscosity & $\mu$ & $\num{6.97e-04}$ & $\mathrm{Pa\cdot s}$ & \cite{Bloomfield1998EffectsFluid} \\
        Cardiac frequency & $f$ & 2.22 & Hz &  \cite{Olstad2019CiliaryDevelopment}\\
        Angular cardiac frequency & $\omega = 2\pi f$ & 6.97 & rad/s & \cite{Olstad2019CiliaryDevelopment}\\
        Cilia tangential traction & $\tau$ & \num{6.5e-04} & Pa & * \\
        Pressure amplitude & $A$ & \num{1.5e-03} & Pa & * \\
        %Ref. frame motion amplitude & $A_{\mathrm{ref}}$ & \num{9.19e-04} & m/s & *\\
        Exosome diffusion coefficient & $D_1$ & $\num{1.63e-12}$ & $\mathrm{m^2/s}$ & \textdagger\\
        STM+GFP diffusion coefficient  & $D_2$ & $\num{5.75e-11}$ & $\mathrm{m^2/s}$ & \textdagger\\
        Dendra2 diffusion coefficient  & $D_3$ & $\num{1.15e-10}$ & $\mathrm{m^2/s}$ & \textdagger\\
    \bottomrule
    \end{tabular}
\end{table}

@@

\subsection*{Model versions and physiological scenarios}\label{subsec:model_versions}
We have introduced boundary conditions to model two separate driving forces of cerebrospinal fluid (CSF) flow, namely motile cilia and cardiac pulsations. To distinguish the impact of each component, three different setups for the flow model were considered~(\cref{tab:flow_model_versions}): 
\begin{itemize}
    \item Model 0: The baseline model representing a wild-type zebrafish. Both cilia forces and cardiac pulsations are included in the model.

@@

    %\item Model IIb: Only cardiac pulsations are modeled, using the domain movement approach. No cilia forces.
\end{itemize}
\begin{table}[h!]
    \centering
    \caption{A summary of the flow model versions. For all model versions, we consider a free slip boundary condition on $\Gs=\Gamma\setminus\left(\Gc\cup\Gp\right)$.}
    \label{tab:flow_model_versions}
    \begin{tabular}{c|c@{\hskip 0.4in}c}
       \toprule
       Model & Driving force & Flow BCs \\ %& Domain motion \\
       \midrule
       Baseline (0)   & Cilia and cardiac & \makecell[l]{Impermeability $\uu\cdot\nn=0$ and tangential traction $\bsigpar=\btau$ on $\Gc$, \\ normal pressure $\Tilde{p}(t)$ on $\Gp$} \\[0.3in] %& No\\[0.3in]
       No cardiac (I) & Cilia   & \makecell[l]{Impermeability $\uu\cdot\nn=0$ and tangential traction $\bsigpar=\btau$ on $\Gc$} \\[0.3in]%& No\\[0.3in]
       No cilia (II) & Cardiac & \makecell[c]{Normal pressure $\Tilde{p}(t)$ on $\Gp$} \\% & No\\[0.3in]
       %No cilia (IIb) & Cardiac & \makecell[l]{Impermeability $\mathbf{w}\cdot\nn=0$ and \\ free slip $\bsigpar(\mathbf{w}, p)=\mathbf{0}$ on $\Gamma$}  & Yes \\
       \bottomrule
    \end{tabular}
\end{table}

When simulating transport of solute in the ventricles, we consider the same six different regions of interest that were considered in the photoconversion experiments. In region number one (the photoconversion region), we use the function~\eqref{eq:injection_curve} to enforce the concentration such that it represents the experimental data of photoconversion of proteins. We consider the three different diffusion coefficients introduced in previous sections to simulate and compare results for transport of molecules of varying sizes.

To simulate transport in the pathological case of region-specific cilia absence, we consider three scenarios where we modify the boundary conditions of the CSF flow model. The cilia tangential stress boundary condition for the boundary where we remove cilia is changed to a free-slip boundary condition, prescribing zero tangential forces to represent absence of cilia motion in the region where we consider cilia removal. We consider removal of cilia in the three separate cilia-covered regions of the telencephalic (anterior) ventricle, the dorsal part of the diencephalic (middle) ventricle, and the ventral part of the diencephalic ventricle~(\cref{fig:figure1}, panel~\textbf{d}). For the three scenarios, we simulate CSF flow modeled by the Stokes equations~\eqref{eq:stokes_eqs} with the modified boundary conditions, and use the resulting velocity field in transport simulations.

The brain ventricular geometry varies significantly across zebrafish individuals, both healthy and sick animals. Motivated by an interest in how these variations can impact solute transport in the brain ventricles, we study the impact on simulation results when modifying the computational geometry and mesh in four different scenarios. Because our computational mesh is inflated by the dye-injection performed before image segmentation, we study four different mesh versions where we shrink: the fore-mid brain connection; the middle ventricle; the mid-hind brain connection; and all of the three above simultaenously. We used Blender~\cite{Community2018BlenderPackage} to modify the brain ventricles geometry.

\subsection*{Quantities of interest}
We quantify the transport of solute by calculating the mean concentrations
\begin{equation}
    \overline{c}_i(t) = \frac{1}{|\Omega_i|}\int_{\Omega_i} c(\xx, t)\,\dx,
    \label{eq:c_mean_i}
\end{equation}
in the regions of interest (ROIs) $\Omega_i$ as functions of time, where $|\Omega_i|$ is the volume of the domain $\Omega_i$. The dynamics of ROI 1 follows from imposing~\cref{eq:injection_curve}. As another way of quantifying transport, we report the times $\hat{t}_i$ when the mean concentrations $\overline{c}_i$ first exceed a threshold value $\hat{c}_i$. We use $\hat{c}_i=0.25$ for regions of interest 1--4 and $\hat{c}_i=0.10$ for regions 5 and 6.

@@

In the absence of pressure boundary conditions, the pressure in the Stokes equations is only determined up to an additive constant. Thus, with the boundary conditions used in model version I, the pressure is not uniquely defined and discretization of the Stokes equations leads to a singular linear equation system. To make the discrete problem well-posed for this setup, the right-hand side vector of the linear system assembled after discretization of the Stokes equations equations is orthogonalized with respect to the basis of the nullspace of the pressure to ensure existence of solutions. This nullspace is the set of (real) scalar constants. Furthermore, information of the pressure nullspace is provided to the linear solver, such that a unique solution is determined.

@@

%%--------------RESULTS--------------%%
\section*{Results}\label{sec:results}
\subsection*{Cerebrospinal fluid flow in the brain ventricles is compartmentalized}
With cilia forces and cardiac forces represented in the flow model (model version 0), we observe a vortex-structured flow pattern in the anterior and middle ventricles~(\cref{fig:figure1}, panel \textbf{e}). This rotational motion shows that cilia-mediated flow is dominating in these regions, as the vortex structures are a consequence of the applied tangential stresses on the ciliated boundaries combined with the assumption of fluid incompressibility. In the posterior ventricle, we observe directional flow as the pressure-driven flow is dominating in this region. The cardiac pressure forces induce a pulsatile motion of the cerebrospinal fluid synchronized with the cardiac cycle, with velocity magnitudes of the pulsatile motion being approximately an order of magnitude smaller than the velocities related to the cilia-driven motion. Over a cardiac cycle, the maximum velocity magnitude in the ventricles was 27.7 $\mu$m/s, located dorsally in the posterior part of the middle ventricle. This maximum velocity was attained at the time instant coinciding with the peak of the cardiac pulsatile flow flowing caudally, as a result of both the tangential stress boundary condition and the pulsatile pressure-driven flow being aligned with the rostrocaudal axis (the $x$ axis). A calculation using the observed maximum velocity of 27.7 $\mu$m/s, and choosing the approximate height of the middle ventricle $110 \ \mu$m as a characteristic length scale, yields a Reynolds number $\mathrm{Re} \approx 0.004$, justifying the assumption of Stokes flow.

\subsection*{Vortex structures vanish in the absence of cilia}
The pulsatile motion is absent when disregarding cardiac motion~(model I, \cref{fig:figure1}, panel~\textbf{f}). For model I, the boundary conditions are not time-dependent, rendering the solution to the Stokes equations~\eqref{eq:stokes_eqs} a steady-state flow field. Streamlines demonstrate the compartmentalization of the cerebrospinal fluid flow driven by cilia, and we observe highest vorticity in the vicinity of cilia. Excluding the cilia tangential traction from the model (model II), vortex-structured compartmentalization of the ventricular flow does not occur, and the flow is directional and pulsatile as a result of the pressure boundary conditions imposed~(\cref{fig:figure1}, panel \textbf{g}).

@@

Transport of solute is quantified by the profile of the mean concentrations $\overline{c}_i(t)$ in the regions of interest (ROIs) $\Omega_i$ as functions of time~(\cref{fig:figure2_compare_diffusion_coefficients}, panels \textbf{b--d}).  An illustration of the ROIs is provided in \cref{fig:figure2_compare_diffusion_coefficients}, panel~\textbf{e}. The dynamics of ROI 1 follow from imposing~\cref{eq:injection_curve}. For ROIs 2--6, we observe more rapid concentration increases when the diffusion coefficient increases.  For all three diffusion coefficients, the dynamics of ROI 6 is the slowest. Region 5 exhibits a more rapid increase in mean concentration, and an even faster increase is observed for ROIs 2--4. The profiles of the latter three regions are similar, displaying only small differences in all of the three plots. \lyng{HH: Add some quantification in terms of percentages, e.g. final values of $\overline{c}_i$ for the different $D_i$?}
%\lyng{HH: is this really true?: Note that the total mass added to the system is not exactly the same for all three diffusion coefficients, since the total mass added depends on how fast the mass added in ROI 1 is transported away from that region.}

We report the times $\hat{t}$ when the mean concentrations $\overline{c}_i$ first exceed a threshold value $\hat{c}$~(\cref{fig:figure2_compare_diffusion_coefficients}, panel \textbf{f}). Recall that $\hat{c}=0.25$ for regions of interest 1--4 and $\hat{c}=0.10$ for regions 5 and 6. The choice of different threshold values is because of the slower dynamics in ROIs 5 and 6. In region 6, the threshold is never exceeded for $D_1$, with the mean concentration close to zero at the end of the simulation.

Finally, we calculate Péclet numbers for the entire ventricular geometry using estimates of the diffusion and advection timescales $t_d=L_p^2/D$ and $t_a=L_p/U_p$. Recall that
\begin{equation*}
    \mathrm{Pe} = \frac{\mathrm{diffusion \ timescale}}{\mathrm{advection \ timescale}} = \frac{t_d}{t_a},
\end{equation*}
so that if $\mathrm{Pe} > 1$, transport is dominated by advection, and if $\mathrm{Pe} < 1$ transport is dominated by diffusion. For the length scale $L_p$ we choose the approximate length of the ventricles geometry in the $x$-direction, $L_p=600 \ \mu$m. We use a mean velocity $\overline{U}$ as the characteristic velocity $U_p$. The mean velocity $\overline{U}$ is calculated by first averaging the velocity in space:
\begin{equation*}
    \overline{u}^2 = \frac{1}{|\Omega|}\int_{\Omega}\uu\cdot\uu\dx,
\end{equation*}
and then averaging $\overline{u}$ in time over one cardiac cycle to obtain $\overline{U}=1.41 \ \mu$m/s. This yields the three Péclet numbers $\mathrm{Pe}_1=519$, $\mathrm{Pe}_2=14.7$ and $\mathrm{Pe}_3=7.4$ for $D_1$, $D_2$ and $D_3$, respectively. \lyng{HH: move choice of length scale/calculation of velocity scale to Methods? Same question for Reynolds number calculation.}

\subsection*{Comparing experimental and computational results}
Transport of photoconverted proteins is quantified by the change in fluorescence intensity $\Delta F=(F(t)-F_0)/F_0$ over time~(\cref{fig:figure3_compare_exp_and_sim_control}, panel \textbf{a}). The solid lines are the mean values of $\Delta F$ for the fish cohort, with shaded regions covering one standard deviation. The first region of interest (ROI 1) is the laser-targeted region. Qualitatively, the fluorescence intensity change in this region resembles intensity dynamics observed in experiments that studied the photoconversion properties of Dendra2 fluorescent proteins~\cite{Makarov2014Steady-stateDendra2}. The transport to the other regions can be separated into two groups: the fluorescence intensity change $\Delta F$ for ROIs 2--4 in the middle ventricle are all similar, whereas the more distal regions ROI 5 and 6 (anterior and posterior ventricle, respectively) exhibit a slower increase in $\Delta F$. In ROIs 2--4, the $\Delta F$ curves are concave, similar to ROI 1, while in ROIs 5 and 6 the curves are convex. Although the fluorescence intensity increases throughout all of the experiments, the maximum $\Delta F$ is not necessarily attained at the final time of the experiments. At the final time, the mean fluoresence intensities in regions 2--4 all exceed 0.60 (0.63, 0.71, 0.67, respectively) of the final value $F_0$ in ROI 1, wheras ROIs 5 and 6 reach mean values of 0.29 and 0.18, respectively.
